---
title: "Assignment3"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

## attach all the necessary packages
```{r}
library(tidyverse)
library(gtsummary)
library(tableone)
library(broom)
library(car)
```

## Read data set
```{r}
pisa2009 <- read.csv("C:/Users/rejalu1/OneDrive - Henry Ford Health System/DSC423/DataSets/pisa2009.csv")
summary(pisa2009)
```

## view data
```{r}
head(pisa2009) 
tail(pisa2009)
```

## data structure
```{r}
str(pisa2009)
```

## select variables of interest
```{r}
pisa_select <- pisa2009 %>%
  select(grade = grade
         , male = male
         , raceeth = raceeth
         , preschool = preschool
         , expectbachelors = expectBachelors
         , motherhs = motherHS
         , motherbachelors = motherBachelors
         , motherwork = motherWork
         , fatherhs = fatherHS
         , fatherbachelors = fatherBachelors
         , fatherwork = fatherWork
         , selfbornus = selfBornUS
         , motherbornus = motherBornUS
         , fatherbornus = fatherBornUS
         , englishathome = englishAtHome
         , computerforschoolwork = computerForSchoolwork
         , read30minsaday = read30MinsADay
         , minutesperweekenglish = minutesPerWeekEnglish
         , studentsinenglish = studentsInEnglish
         , schoolhaslibrary = schoolHasLibrary
         , publicschool = publicSchool
         , urban = urban
         , schoolsize = schoolSize
         , readingscore = readingScore
        )
```

## checking for missing values
```{r}
sum(is.na(pisa_select))
```



## data cleaning
```{r}
pisa_clean <- pisa_select %>%
  transmute(grade = as.factor(grade)
            , male = male
            , raceeth = as.factor(raceeth)
            , preschool = preschool
            , expectbachelors = expectbachelors
            , motherhs = motherhs
            , motherbachelors = motherbachelors
            , motherwork = motherwork
            , fatherhs = fatherhs
            , fatherbachelors = fatherbachelors
            , fatherwork = fatherwork
            , selfbornus = selfbornus
            , motherbornus = motherbornus
            , fatherbornus = fatherbornus
            , englishathome = englishathome
            , computerforschoolwork = computerforschoolwork
            , read30minsaday = read30minsaday
            , minutesperweekenglish = minutesperweekenglish
            , studentsinenglish = studentsinenglish
            , schoolhaslibrary = schoolhaslibrary
            , publicschool = publicschool
            , urban = urban
            , schoolsize = schoolsize
            , readingscore = readingscore) %>% 
  
mutate(raceeth = relevel(raceeth, ref = 'White')
       , grade = relevel(grade, ref = '10')
       )

## we create a matrix for raceeth
raceethdummies.matrix <- model.matrix(~pisa_clean$raceeth)
## convert the model matrix into a data frame
raceethdummies.frame <- data.frame(raceethdummies.matrix)
pisa_clean <- cbind(pisa_clean, raceethdummies.frame)

##Create a matrix for grade
gradedummies.matrix <- model.matrix(~pisa_clean$grade)
## convert the model matrix into a data frame
gradedummies.frame <- data.frame(gradedummies.matrix)
pisa_clean <- cbind(pisa_clean, gradedummies.frame)
```


## rename and select variables of interest after data cleaning
```{r}
pisacleansed <- pisa_clean %>%
  select(readingscore = readingscore
            , male = male
            , preschool = preschool
            , expectbachelors = expectbachelors
            , motherhs = motherhs
            , motherbachelors = motherbachelors
            , motherwork = motherwork
            , fatherhs = fatherhs
            , fatherbachelors = fatherbachelors
            , fatherwork = fatherwork
            , selfbornus = selfbornus
            , motherbornus = motherbornus
            , fatherbornus = fatherbornus
            , englishathome = englishathome
            , computerforschoolwork = computerforschoolwork
            , read30minsaday = read30minsaday
            , minutesperweekenglish = minutesperweekenglish
            , studentsinenglish = studentsinenglish
            , schoolhaslibrary = schoolhaslibrary
            , publicschool = publicschool
            , urban = urban
            , schoolsize = schoolsize
            , american_indian_alaska_native = pisa_clean.raceethAmerican.Indian.Alaska.Native
            , asian = pisa_clean.raceethAsian
            , black = pisa_clean.raceethBlack
            , hispanic = pisa_clean.raceethHispanic
            , morethanonerace = pisa_clean.raceethMore.than.one.race
            , nativehawaiianOtherPacificIslander = pisa_clean.raceethNative.Hawaiian.Other.Pacific.Islander
            , G8 = pisa_clean.grade8
            , G9 = pisa_clean.grade9
            , G11 = pisa_clean.grade11
            , G12 = pisa_clean.grade12)
```



## Perform an eighty-twenty split to partition the data set
## into train and test data sets.
```{r}
set.seed(123)
partition <- sample(2,nrow(pisacleansed),replace = TRUE, prob= c(0.80,0.20))
train <- pisacleansed[partition==1, ]
test <-  pisacleansed[partition==2, ]
```

## check for multicollinearity
```{r}
cor(pisacleansed)
```


## full model
## Adjusted R-squared:  0.302 for A and correlation of 0.576
```{r}
modelA <- lm(readingscore ~ ., data = train)
summary(modelA)
prediction <- predict(modelA,test)
actual =test$readingscore
cor(prediction, actual)
vif(modelA)
```


## Pruned selfbornus
## Adjusted R-squared:  0.302 and correlation of 0.576
```{r}
modelB <- lm(readingscore ~ male+preschool+expectbachelors+motherhs+motherbachelors+motherwork+fatherhs+fatherbachelors
+ fatherwork + motherbornus+fatherbornus+englishathome+computerforschoolwork+read30minsaday
+ minutesperweekenglish+studentsinenglish+schoolhaslibrary+publicschool+urban+schoolsize
+ american_indian_alaska_native+asian+black+hispanic+morethanonerace+nativehawaiianOtherPacificIslander
+ G8 + G9 +G11 + G12, data = train)
summary(modelB)
prediction <- predict(modelB,test)
actual =test$readingscore
cor(prediction, actual)
```


## Prune out school has a library
## Adjusted R-squared:  0.303 and correlation of 0.576
```{r}
modelC <- lm(readingscore ~ male+preschool+expectbachelors+motherhs+motherbachelors+motherwork+fatherhs+fatherbachelors
+ fatherwork + motherbornus+fatherbornus+englishathome+computerforschoolwork+read30minsaday
+ minutesperweekenglish+studentsinenglish + publicschool+urban+schoolsize
+ american_indian_alaska_native+asian+black+hispanic+morethanonerace+nativehawaiianOtherPacificIslander
+ G8 + G9 +G11 + G12, data = train)
summary(modelC)
prediction <- predict(modelC,test)
actual =test$readingscore
cor(prediction, actual)
```


## Prune out urban
## Adjusted R-squared:  0.303  and correlation of 0.576
```{r}
modelD <- lm(readingscore ~ male + preschool + expectbachelors + motherhs + motherbachelors + motherwork + fatherhs + fatherbachelors + fatherwork + motherbornus + fatherbornus + englishathome + computerforschoolwork + read30minsaday + minutesperweekenglish+studentsinenglish + publicschool + schoolsize + american_indian_alaska_native + asian + black + hispanic + morethanonerace + nativehawaiianOtherPacificIslander + G8 + G9 +G11 + G12, data = train)
summary(modelD)
prediction <- predict(modelD,test)
actual =test$readingscore
cor(prediction, actual)
```

## Pruned studentsinenglish
## Adjusted R-squared:  0.303 and correlation of 0.576
```{r}
modelE <- lm(readingscore ~ male + preschool + expectbachelors + motherhs + motherbachelors + motherwork + fatherhs + fatherbachelors + fatherwork + motherbornus + fatherbornus + englishathome + computerforschoolwork + read30minsaday + minutesperweekenglish + publicschool + schoolsize + american_indian_alaska_native + asian + black + hispanic + morethanonerace + nativehawaiianOtherPacificIslander + G8 + G9 +G11 + G12, data = train)
summary(modelE)
prediction <- predict(modelE,test)
actual =test$readingscore
cor(prediction, actual)
```

## Pruned Preschool
Adjusted R-squared:  0.303  and correlation of 0.576
```{r}
modelF <- lm(readingscore ~ male  + expectbachelors + motherhs + motherbachelors + motherwork + fatherhs + fatherbachelors + fatherwork + motherbornus + fatherbornus + englishathome + computerforschoolwork + read30minsaday + minutesperweekenglish + publicschool + schoolsize + american_indian_alaska_native + asian + black + hispanic + morethanonerace + nativehawaiianOtherPacificIslander + G8 + G9 +G11 + G12, data = train)
summary(modelF)
prediction <- predict(modelF,test)
actual =test$readingscore
cor(prediction, actual)
```

## Pruned fatherbornus
## Adjusted R-squared:  0.304 and correlation of 0.576
```{r}
modelG <- lm(readingscore ~ male  + expectbachelors + motherhs + motherbachelors + motherwork + fatherhs + fatherbachelors + fatherwork + motherbornus  + englishathome + computerforschoolwork + read30minsaday + minutesperweekenglish + publicschool + schoolsize + american_indian_alaska_native + asian + black + hispanic + morethanonerace + nativehawaiianOtherPacificIslander + G8 + G9 +G11 + G12, data = train)
summary(modelG)
prediction <- predict(modelG,test)
actual =test$readingscore
cor(prediction, actual)
```

## pruned fatherwork
## Adjusted R-squared:  0.304 and correlation of 0.576
```{r}
modelH <- lm(readingscore ~ male  + expectbachelors + motherhs + motherbachelors + motherwork + fatherhs + fatherbachelors  + motherbornus  + englishathome + computerforschoolwork + read30minsaday + minutesperweekenglish + publicschool + schoolsize + american_indian_alaska_native + asian + black + hispanic + morethanonerace + nativehawaiianOtherPacificIslander + G8 + G9 +G11 + G12, data = train)
summary(modelH)
prediction <- predict(modelH,test)
actual =test$readingscore
cor(prediction, actual)
```

## pruned motherhs
## Adjusted R-squared:  0.304 and correlation of 0.576
```{r}
modeli <- lm(readingscore ~ male  + expectbachelors  + motherbachelors + motherwork + fatherhs + fatherbachelors  + motherbornus  + englishathome + computerforschoolwork + read30minsaday + minutesperweekenglish + publicschool + schoolsize + american_indian_alaska_native + asian + black + hispanic + morethanonerace + nativehawaiianOtherPacificIslander + G8 + G9 +G11 + G12, data = train)
summary(modeli)
prediction <- predict(modeli,test)
actual =test$readingscore
cor(prediction, actual)
```

## Pruned motherbornus
## Adjusted R-squared:  0.304 and 0.575
```{r}
modelk <- lm(readingscore ~ male  + expectbachelors  + motherbachelors + motherwork + fatherhs + fatherbachelors  + englishathome + computerforschoolwork + read30minsaday + minutesperweekenglish + publicschool + schoolsize + american_indian_alaska_native + asian + black + hispanic + morethanonerace + nativehawaiianOtherPacificIslander + G8 + G9 +G11 + G12, data = train)
summary(modelk)
prediction <- predict(modelk,test)
actual =test$readingscore
cor(prediction, actual)
```

## Prunned  motherwork 
```{r}
modell <- lm(readingscore ~ male  + expectbachelors  + motherbachelors  + fatherhs + fatherbachelors  + englishathome + computerforschoolwork + read30minsaday + minutesperweekenglish + publicschool + schoolsize + american_indian_alaska_native + asian + black + hispanic + morethanonerace + nativehawaiianOtherPacificIslander + G8 + G9 +G11 + G12, data = pisacleansed)
summary(modell)
prediction <- predict(modelk,test)
actual =test$readingscore
cor(prediction, actual)
```


## 5 fold cross validation to validate  modelA
## Average mean square error is 210964 
```{r}
library(DAAG)
outA <- cv.lm(data = test
             , form.lm = formula((readingscore ~ .))
             , plotit = "Observed", m=5)
```

## 5 fold cross validation to validate our modelB
## Average mean square error is 5741 
```{r}
library(DAAG)
outB <- cv.lm(data = test
             , form.lm = formula((readingscore ~ male + preschool + expectbachelors + motherhs + motherbachelors + motherwork + fatherhs+fatherbachelors
+ fatherwork + motherbornus + fatherbornus + englishathome + computerforschoolwork + read30minsaday
+ minutesperweekenglish + studentsinenglish + schoolhaslibrary + publicschool + urban + schoolsize
+ american_indian_alaska_native + asian + black + hispanic + morethanonerace + nativehawaiianOtherPacificIslander + G8 + G9 +G11 + G12))
             , plotit = "Observed", m=5)
```

## 5 fold cross validation to validate our modelC
## Average mean square error is 5709 
```{r}
library(DAAG)
outC <- cv.lm(data = test
             , form.lm = formula((readingscore ~ male + preschool + expectbachelors + motherhs + motherbachelors + motherwork + fatherhs + fatherbachelors
+ fatherwork + motherbornus + fatherbornus + englishathome + computerforschoolwork + read30minsaday
+ minutesperweekenglish + studentsinenglish + publicschool + urban + schoolsize
+ american_indian_alaska_native + asian + black + hispanic + morethanonerace + nativehawaiianOtherPacificIslander + G8 + G9 +G11 + G12))
             , plotit = "Observed", m=5)
```

## 5 fold cross validation to validate our modelD
## Average mean square error is 5668 
```{r}
library(DAAG)
outD <- cv.lm(data = test
             , form.lm = formula((readingscore ~ male + preschool + expectbachelors + motherhs + motherbachelors + motherwork + fatherhs + fatherbachelors + fatherwork + motherbornus + fatherbornus + englishathome + computerforschoolwork + read30minsaday + minutesperweekenglish+studentsinenglish + publicschool + schoolsize + american_indian_alaska_native + asian + black + hispanic + morethanonerace + nativehawaiianOtherPacificIslander + G8 + G9 +G11 + G12))
             , plotit = "Observed", m=5)
```

## 5 fold cross validation to validate our modelE
## Average mean square error is 5692
```{r}
outE <- cv.lm(data = test
             , form.lm = formula((readingscore ~ male + preschool + expectbachelors + motherhs + motherbachelors + motherwork + fatherhs + fatherbachelors + fatherwork + motherbornus + fatherbornus + englishathome + computerforschoolwork + read30minsaday + minutesperweekenglish + publicschool + schoolsize + american_indian_alaska_native + asian + black + hispanic + morethanonerace + nativehawaiianOtherPacificIslander + G8 + G9 +G11 + G12))
             , plotit = "Observed", m=5)
```

## 5 fold cross validation to validate our modelF
## Average mean square error of 5690 
```{r}
outF <- cv.lm(data = test
             , form.lm = formula((readingscore ~ male  + expectbachelors + motherhs + motherbachelors + motherwork + fatherhs + fatherbachelors + fatherwork + motherbornus + fatherbornus + englishathome + computerforschoolwork + read30minsaday + minutesperweekenglish + publicschool + schoolsize + american_indian_alaska_native + asian + black + hispanic + morethanonerace + nativehawaiianOtherPacificIslander + G8 + G9 +G11 + G12))
             , plotit = "Observed", m=5)
```

## 5 fold cross validation to validate our modelG
## Average mean square error of 5648
```{r}
outF <- cv.lm(data = test
             , form.lm = formula((readingscore ~ male  + expectbachelors + motherhs + motherbachelors + motherwork + fatherhs + fatherbachelors + fatherwork + motherbornus  + englishathome + computerforschoolwork + read30minsaday + minutesperweekenglish + publicschool + schoolsize + american_indian_alaska_native + asian + black + hispanic + morethanonerace + nativehawaiianOtherPacificIslander + G8 + G9 +G11 + G12))
             , plotit = "Observed", m=5)
```

## 5 fold cross validation to validate our modelk
## Average mean square error of 
```{r}
outK <- cv.lm(data = test
             , form.lm = formula((readingscore ~ male  + expectbachelors  + motherbachelors  + fatherhs + fatherbachelors  + englishathome + computerforschoolwork + read30minsaday + minutesperweekenglish + publicschool + schoolsize + american_indian_alaska_native + asian + black + hispanic + morethanonerace + nativehawaiianOtherPacificIslander + G8 + G9 +G11 + G12))
             , plotit = "Observed", m=5)
```

## table1
## Shows descriptive statistics about your data
```{r}
CreateTableOne(vars = colnames(pisacleansed)
               , data = pisacleansed)
```

## Count the number of times a value appears in the grade variable
```{r}
grade <- table(pisacleansed$nativehawaiianOtherPacificIslander)
grade
```



## model selection for second oder terms
## Adjusted R-squared:  0.316 
```{r}
pisacleansed$minutesperweekenglishSQ <- pisacleansed$minutesperweekenglish^2
modeli <- lm(readingscore ~ male  + expectbachelors  + motherbachelors + motherwork + fatherhs + fatherbachelors  + motherbornus  + englishathome + computerforschoolwork + read30minsaday + minutesperweekenglish + publicschool + schoolsize + american_indian_alaska_native + asian + black + hispanic + morethanonerace + nativehawaiianOtherPacificIslander + G8 + G9 +G11 + G12 + minutesperweekenglishSQ, data = pisacleansed)
summary(modeli)
```


## Prune motherwork 
## Adjusted R-squared:  0.316 
```{r}
pisacleansed$minutesperweekenglishSQ <- pisacleansed$minutesperweekenglish^2
modeli <- lm(readingscore ~ male  + expectbachelors  + motherbachelors  + fatherhs + fatherbachelors  + motherbornus  + englishathome + computerforschoolwork + read30minsaday + minutesperweekenglish + publicschool + schoolsize + american_indian_alaska_native + asian + black + hispanic + morethanonerace + nativehawaiianOtherPacificIslander + G8 + G9 +G11 + G12 + minutesperweekenglishSQ, data = pisacleansed)
summary(modeli)
```


##Prune asian
## Adjusted R-squared:  0.316 
```{r}
pisacleansed$minutesperweekenglishSQ <- pisacleansed$minutesperweekenglish^2
modeli <- lm(readingscore ~ male  + expectbachelors  + motherbachelors  + fatherhs + fatherbachelors  + motherbornus  + englishathome + computerforschoolwork + read30minsaday + minutesperweekenglish + publicschool + schoolsize + american_indian_alaska_native  + black + hispanic + morethanonerace + nativehawaiianOtherPacificIslander + G8 + G9 +G11 + G12 + minutesperweekenglishSQ, data = pisacleansed)
summary(modeli)
```


## Pruned nativehawaiianOtherPacificIslander
## Adjusted R-squared:  0.316
```{r}
pisacleansed$minutesperweekenglishSQ <- pisacleansed$minutesperweekenglish^2
modeli <- lm(readingscore ~ male  + expectbachelors  + motherbachelors  + fatherhs + fatherbachelors  + motherbornus  + englishathome + computerforschoolwork + read30minsaday + minutesperweekenglish + publicschool + schoolsize + american_indian_alaska_native  + black + hispanic + morethanonerace  + G8 + G9 +G11 + G12 + minutesperweekenglishSQ, data = pisacleansed)
summary(modeli)
```

## Pruned motherbornus
## Adjusted R-squared:  0.316
```{r}
pisacleansed$minutesperweekenglishSQ <- pisacleansed$minutesperweekenglish^2
modeli <- lm(readingscore ~ male  + expectbachelors  + motherbachelors  + fatherhs + fatherbachelors    + englishathome + computerforschoolwork + read30minsaday + minutesperweekenglish + publicschool + schoolsize + american_indian_alaska_native  + black + hispanic + morethanonerace  + G8 + G9 +G11 + G12 + minutesperweekenglishSQ, data = pisacleansed)
summary(modeli)
```

## Prune fatherhs
## Adjusted R-squared:  0.315 
```{r}
pisacleansed$minutesperweekenglishSQ <- pisacleansed$minutesperweekenglish^2
modeli <- lm(readingscore ~ male  + expectbachelors  + motherbachelors   + fatherbachelors    + englishathome + computerforschoolwork + read30minsaday + minutesperweekenglish + publicschool + schoolsize + american_indian_alaska_native  + black + hispanic + morethanonerace  + G8 + G9 +G11 + G12 + minutesperweekenglishSQ, data = pisacleansed)
summary(modeli)
```



## Check for interaction terms 
## Adjusted R-squared:  0.317
## two interactio terms
```{r}
pisacleansed$minutesperweekenglishSQ <- pisacleansed$minutesperweekenglish^2
modeli <- lm(readingscore ~ male  + expectbachelors  + motherbachelors   + fatherbachelors    + englishathome + computerforschoolwork + read30minsaday + minutesperweekenglish + publicschool + schoolsize + american_indian_alaska_native  + black + hispanic + morethanonerace  + G8 + G9 +G11 + G12 + minutesperweekenglishSQ + read30minsaday *minutesperweekenglish + motherbachelors* fatherbachelors, data = pisacleansed)
summary(modeli)
plot(modeli)
```


## Removing specific rows in r
```{r}
## Adding an index column to my data set
pisacleansed$generated_uid <- 1:nrow(pisacleansed)
pisacleansed <- pisacleansed[-c(2657,507,1879),]
```

##read30minsaday:minutesperweekenglish
```{r}
pisacleansed$minutesperweekenglishSQ <- pisacleansed$minutesperweekenglish^2
modeli <- lm(readingscore ~ male  + expectbachelors  + motherbachelors   + fatherbachelors    + englishathome + computerforschoolwork + read30minsaday + minutesperweekenglish + publicschool + schoolsize + american_indian_alaska_native  + black + hispanic + morethanonerace  + G8 + G9 +G11 + G12 + minutesperweekenglishSQ  + motherbachelors* fatherbachelors, data = pisacleansed)
summary(modeli)
```


## Variable transformation
## log of both independent variable
## Adjusted R-squared:  0.322
```{r}
pisacleansed$minutesperweekenglishSQ <- pisacleansed$minutesperweekenglish^2
modeli <- lm(log(readingscore + 1) ~ male  + expectbachelors  + motherbachelors   + fatherbachelors    + englishathome + computerforschoolwork + read30minsaday + minutesperweekenglish + publicschool + schoolsize + american_indian_alaska_native  + black + hispanic + morethanonerace  + G8 + G9 +G11 + G12 + minutesperweekenglishSQ  + motherbachelors* fatherbachelors, data = pisacleansed)
summary(modeli)
vif(modeli)
```


##model evaluation

## Perform an eighty-twenty split to partition the data set
## into train and test data sets so that the model can be evaluated
```{r}
set.seed(123)
partition <- sample(2,nrow(pisacleansed),replace = TRUE, prob= c(0.80,0.20))
train <- pisacleansed[partition==1, ]
test <-  pisacleansed[partition==2, ]

modeli <- lm(log(readingscore + 1) ~ male  + expectbachelors  + motherbachelors   + fatherbachelors    + englishathome + computerforschoolwork + read30minsaday + minutesperweekenglish + publicschool + schoolsize + american_indian_alaska_native  + black + hispanic + morethanonerace  + G8 + G9 +G11 + G12 + minutesperweekenglishSQ + read30minsaday *minutesperweekenglish + motherbachelors* fatherbachelors, data = train)
summary(modeli)
prediction <- predict(modeli,test)
actual =test$readingscore
cor(prediction, actual)
```




## Creating sample dataset
```{r}
dsample <- train[sample(1:nrow(train), 1000,replace = FALSE),]
##plot(dsample)
```




